---
title: "Lab 1"
author: "Xuyi Ren"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Question 1:**

Create a new R Markdown file and copy the entirety of this question over
to the new file (we will do this for all questions in this lab). Then,
proceed with the instructions below.

Between the stars below, do the following:

1.  Use two \# to create a header that says About Me

    ## About Me

2.  Type your first name in bold and your last name in italics **Xuyi**
    *Ren*

3.  Create a bullet point list of the people sitting on either side of
    you

-   Xuyi Ren
-   Yingying Li
-   Yingying Li

4.  Create a numbered list of your 3 least favorite animals
5.  snake
6.  fish
7.  lizard
8.  Once you have done this, Knit to PDF by clicking the bar of yarn
    above and verify that everything looks like it should.

Question 2:

Again, copy the entirety of this question into the R Markdown file you
created for Question 1.

Let’s practice creating vectors and subsetting with a short exercise.

1.  First, create an R code chunk between the rows of stars below
    (Ctrl+Alt+I is quick way to do this)

```{r}

```

2.  Next, create a vector called x that has all of the numbers from 11
    to 20

```{r}
x <- c(11:20)
```

3.  Use square brackets and subsetting to select the first five numbers
    of this vector.

```{r}
x[1:5]
```

Question 3:

For this question, we will be using the HappyPlanet data that we have
just looked at:

1.  Part A Copy the code above to read the Happy Planet data into your
    own R Markdown file, saving the dataset to a variable called planet

```{r}
planet <- read.csv("https://collinn.github.io/data/HappyPlanet.csv")
```

2.  Part B Looking at the Happy Planet data, explain in one or two
    sentences what constitutes an observation in this dataset (what is
    the data being recorded from)

It is a dataset of countries and their life expectancy, happiness score,
and ecological footprint.

3.  Part C Using \$ to extract columns from the dataset, find the mean
    life expectancy of all countries in the dataset? (Hint: what
    functions have we seen already in this lab?)

```{r}
mean(planet$LifeExpectancy)
```

4.  Part D Are there any variables in this dataset that are stored as a
    numeric that would be better suited as a categorical variable?
    Explain your answer

Yes, they are: 1. Country 2. Region 3. LifeExpectance 4. Happiness.Score
5. Ecological.Footprint

They would be better suited as a categorical because they are not
numbers.
